Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2020-12-15 manuscript.tex
Whitespace changes [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/eff43d9d3566f72329894e11e468eccd5bac3323.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/eff43d9d3566f72329894e11e468eccd5bac3323/checks
https://github.com/habi/acinar-analysis-manuscript/runs/423752790

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681} and is under revision in the \href{https://journals.physiology.org/journal/ajplung}{\emph{American Journal of Physiology - Lung Cellular and Molecular Physiology}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/d6fc17bd66bdca5cbff30d53e1dbc02d5351609b...d0fe50ee2f2a7caa83758cf8364e3629637ef31a}{has been updated} since the first round of review.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/d0fe50ee2f2a7caa83758cf8364e3629637ef31a/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/d0fe50ee2f2a7caa83758cf8364e3629637ef31a}{habi/acinar-analysis-manuscript@d0fe50e}
on December 15, 2020.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(d0fe50e)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(d0fe50e)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Switzerland
\end{itemize}

@@

Lung development starts with the formation of the two lung buds.
During the embryonic stage (prenatal days E11--E13 in rats) the major airways and the pleura are formed.
During the pseudoglandular stage (E13--E18.5) most of the remaining airway generations are formed by branching morphogenesis.
Only few additional branches are added during the later stages (canalicular (E18.5--E20) and (E20 to postnatal day P4))~(\protect\hyperlink{ref-TsNwin2E}{34}).
Epithelial differentiation during the canalicular stage and angiogenetic activation of the capillaries lead to the first functional air-blood barriers in the lung~(\protect\hyperlink{ref-ODQCVC23}{33}).
The saccular stage represents an intermediate stage where branching morphogenesis ceases and the developmental program switches to the phase of alveolarization.
During alveolarization, 90~\% of the adult gas-exchange area is formed by lifting off new septa from the existing gas-exchange surface~(\protect\hyperlink{ref-auNiE56G}{5}, \protect\hyperlink{ref-1AhvLCPky}{19}, \protect\hyperlink{ref-XQLPrPTJ}{32}).
Alveolarization itself is divided into two distinct phases, the so-called classical (P4--P21) and continued alveolarization (P14 to approximately P60)~(\protect\hyperlink{ref-d1YV8oOK}{35}, \protect\hyperlink{ref-wnl86DEM}{46}).
During classical alveolarization, new septa are formed starting from preexisting immature septa which contain a double-layered capillary network.
During continued alveolarization, new septa are formed from preexisting mature septa.
In general, during alveolarization the structure of the alveolar septa is changed to increase the efficiency of the lung (\protect\hyperlink{ref-d1YV8oOK}{35}, \protect\hyperlink{ref-wnl86DEM}{46}).

In this manuscript, the timing of lung development is given for rats while the same stages and developmental mechanisms are observed in every placentalian species studied so far, including humans.
However, the timing of lung development and especially the time point of birth relative to lung development is adapted in each species~(\protect\hyperlink{ref-XQLPrPTJ}{32}, \protect\hyperlink{ref-ODQCVC23}{33}, \protect\hyperlink{ref-d1YV8oOK}{35}).

@@

Pulmonary acini represent the gas-exchanging units of the bronchial tree and are defined as the airways distal of the terminal bronchioles~(\protect\hyperlink{ref-BbEFwEVD}{13}).
In humans, the acini contain approximately four generations of respiratory bronchioles before reaching the alveolar ducts.
The alveolar ducts start at the so-called bronchioalveolar duct junction (BADJ) where the lining of the inner airway surface abruptly changes from the cuboidal epithelium of the bronchioles into type 1 and type 2 alveolar epithelium which covers the alveoli.
The small tree of airways distal of the BADJ is called a ventilatory unit~(\protect\hyperlink{ref-WDrIHn9p}{39}).
Since murine lungs do not possess respiratory bronchioles, the murine acini consist of one single ventilatory unit~(\protect\hyperlink{ref-WDrIHn9p}{39}).
In this study we used rat lungs; hence we exclusively speak of acini.
To translate our results from rats to humans, monkeys, or dogs one would have to compare one rat acinus to one single ventilatory unit in these species.

Developmentally, the BADJ is of particular interest.
The junction is formed during the canalicular stage when the epithelia are differentiating.
The generation of the airway in which an individual murine acinus starts is defined during the canalicular stage and remains constant once it is formed.
As a consequence, in rats the number of formed acini also remains constant during the phase of alveolarization and thereafter~(\protect\hyperlink{ref-14OP85b2F}{2}, \protect\hyperlink{ref-uFNlWogb}{3}).
The latter was somehow surprising because in rats the total lung volume increases by roughly a factor of 10 during alveolarization~(\protect\hyperlink{ref-wnl86DEM}{46}).
Therefore, the mean volume of the acini has to increase by approximately the same factor.

The acinar architecture is important for ventilation, airflow~(\protect\hyperlink{ref-CVl41LwO}{8}, \protect\hyperlink{ref-eb0gk6VO}{14}, \protect\hyperlink{ref-1DP2FRUSZ}{16}) and particle deposition~(\protect\hyperlink{ref-18DcNLAv6}{22}, \protect\hyperlink{ref-yHHhvOtP}{36}, \protect\hyperlink{ref-wnl86DEM}{46}).

@@

On single two-dimensional sections---either histological or virtual from tomographic imaging---it is not possible to unambiguously detect which parts of the airways and alveoli are connected three-dimensionally.
To detect and delineate individual acini the full three-dimensional information is needed.
Historically, single or very few acini were delineated and extracted from manually traced serial sections~(\protect\hyperlink{ref-c3aIB7f9}{28}, \protect\hyperlink{ref-CPhMy3Ie}{50}) or sectioned silicone casts~(\protect\hyperlink{ref-1GJECnKBu}{30}) with considerable manual work.
Non-destructive three-dimensional imaging is best suited to acquire data sets that can be used to easily detect, delineate, and study large amounts of single acini.
This has been shown by Vasilescu et al.~(\protect\hyperlink{ref-C6OaY84D}{48}).
In their study, they reconstructed 22 mouse acini from tomographic data sets of four 12 weeks old mice (at least postnatal day 84, fully developed lungs) and compared those with scanning electron microscope images of acinar silicone rubber casts.
Similarly, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) analyzed the volume of 32 mouse acini from six animals, as well as the branching pattern of their internal airways.
All prior work known to us has analyzed considerably fewer acini than the 268 acini presented here and focuses only on a time point, at which the lungs of the animals are already fully developed while our data spans the postnatal days 4 to 60.

@@

While the global volume of the acini is well known and studied, only little is known about the range of the volume of individual acini throughout lung development.
We developed the hereby presented method to analyze large amounts of acini over the course of postnatal lung development in a time-efficient manner since no data on large amounts of acini is available up to now.
Our aim was to understand the size distribution of the acini during lung development including the contribution of the alveoli.
The observed range \emph{de}creased during postnatal lung development by a factor of 6--7, which represents a completely unexpected result.
The obtained data are essential for further investigations, like the influence of the size of the acini on ventilation or gas-washout, respectively.
E.g., we are currently using our computational fluid dynamics model to simulate exactly this influence (own unpublished results and~(\protect\hyperlink{ref-eb0gk6VO}{14})).

@@

In the present study, a superset of the animals described by Haberthür et al.~(\protect\hyperlink{ref-7YLeeyu}{11}) were used.
Tschanz et al.~(\protect\hyperlink{ref-wnl86DEM}{46}) described a biphasic formation of new alveoli on the same set of animals.
The stereological analysis presented here represents a part of the 3R-initiative (replacement, reduction, and refinement)~(\protect\hyperlink{ref-199ALtdJt}{53}) for the ethical use and the reduction of the number of animals sacrificed for science.

Briefly, we extracted the lungs of Sprague-Dawley rats after having deeply anesthetized them with a mixture of medetomidine, midazolam, and fentanyl~(\protect\hyperlink{ref-IdRLVLjl}{7}).
The rats were euthanized by exsanguination during the removal of the lung.
The lungs were fixed with 2.5~\% glutaraldehyde by intratracheal instillation and kept under constant pressure (20 cm H\textsubscript{2}O) during fixation to prevent recoiling of the tissue.
The samples were prepared for tomographic imaging by post-fixation with 1~\% osmium tetroxide and staining with 4~\% uranyl acetate, as well as by dehydration in a graded series of ethanol and embedding in paraffin.
In total we assessed 12 animals, on postnatal days P4, P10, P21, and P60 (N=3 per day.
From now on, we will omit the postnatal prefix and only mention the day.).

@@

Since our samples were substantially larger than the field of view of the detector, we applied the so-called wide-field scanning method~(\protect\hyperlink{ref-VELl7OiR}{12}) to increase the field of view horizontally.
For each sub-scan, we recorded 3578 projections and laterally stitched them so that their combined field of view covered the whole sample width.
Additionally, two or three such wide-field scans were stacked vertically to match the sample height.

@@

The stereological estimation of the alveolar number was performed according to the standards for quantitative assessment of lung structure from the \href{http://www.thoracic.org/}{American Thoracic Society}~(\protect\hyperlink{ref-dNc8FfNn}{17}) to guarantee accurate and unbiased results.
As a basis for the estimation the volumes of the lung lobes were first determined by water displacement (\protect\hyperlink{ref-KGbSQovR}{31}).
The lobe-volume was estimated a second time by the Cavalieri method.
Both volumes were used to calculate the shrinkage, which was used for the correction of the volume data (\protect\hyperlink{ref-wnl86DEM}{46}).

To perform the stereological analysis, each DICOM file from the isolated acini was sampled into a systematic random set of single images using a MATLAB script~(\protect\hyperlink{ref-12Z2YPzm8}{10}).
The stereological assessment was performed with the \href{http://stepanizer.com/}{STEPanizer}~(\protect\hyperlink{ref-nPoQ2EIB}{45}), an easy-to-use software tool for the stereological assessment of digital images.
Details of the stereological assessment have previously been described~(\protect\hyperlink{ref-7YLeeyu}{11}).
Briefly, we counted the appearance or disappearance of alveolar wall interruptions.
Under the assumption that these only occur in the region of the alveolar mouth opening and correspond to alveolar entrance rings~(\protect\hyperlink{ref-QiAxY2i3}{18}) we counted said interruptions on paired images spanning the whole acinus volume.
By using the disector method~(\protect\hyperlink{ref-FJ9FoB4m}{38}) we were thus able to estimate the number of alveoli in 268 of the 701 exported acini.
The acini we analyzed were subsampled from the full data by systematic uniform random sampling~(\protect\hyperlink{ref-dNc8FfNn}{17}).
Briefly, a defined fraction of the acini of one animal were analyzed, e.g.~one third.
During segmentation, the acini were numbered.
Based on these numbers every third acinus was selected, randomly starting with the first, second, or third one.

Previously, we have shown that the stereological assessment of the acinar volume (according to the Cavalieri principle~(\protect\hyperlink{ref-FE9HLB4f}{6})) gives comparable results to an automated volume extraction by counting the segmented pixels in the tomographic data sets~(\protect\hyperlink{ref-7YLeeyu}{11}).
Due to variations in the gray value, the automatic segmentation underestimates the volume of the single acini.
The best-suited approach to obtain unbiased results is to assess the volume manually according to the Cavalieri principle, which is what we did for the 268 acini presented in this manuscript.
The sets of JPG slices and the raw results from the stereological analysis with the STEPanizer are available on request.

@@

All the stereologically assessed data was processed in a \href{http://jupyter.org/}{Jupyter} (\protect\hyperlink{ref-pQ6Wbz73}{21}) notebook, producing all the results, including the statistical data and plots shown below.
The notebook with all its history is freely available on GitHub~(\protect\hyperlink{ref-V87xbt0b}{9}).
The performed calculations are described in detail in the \protect\hyperlink{results}{Results} section.
In the plots, semitransparent circles mark the single observations.
The box shows the 25~\%--75~\% quartiles range of the data.
The whiskers extend to the rest of the distribution.
Outliers are determined as a function of the inter-quartile range and are shown outside the whiskers.
Numerical values in the text are given as averages ± standard deviation.
P-values in the text and figure legends are given as precise numbers, as suggested by Amhrein et al.~(\protect\hyperlink{ref-o21zxPIu}{1}).

Usually, we performed a Shapiro-Wilk test for normality~(\protect\hyperlink{ref-IkHrgIj3}{37}) to test whether we can use an U-test for assessing the differences, namely a two-sided Mann-Whitney rank test~(\protect\hyperlink{ref-MyzxBJ57}{25}).
This rank test was used to assess the differences between the possible combinations.
An additional Kruskal-Wallis H-test for independent samples~(\protect\hyperlink{ref-ON1Bppkk}{23}) was used to test for sample independence.
The statistical analysis was performed in the aforementioned notebook by using the statistical functions of the Scientific Computing Tools for Python~(\protect\hyperlink{ref-8Miti2Gz}{49}) or GraphPad Prism 7.01 (GraphPad Software, San Diego, CA, USA).

@@

The alveolar number shows highly significant differences Šidák-corrected p-value \textless{} 0.00167~(\protect\hyperlink{ref-GtCynqsI}{44})) between all possible combination of days.
All possible combinations of entrance ring counts per day are significantly different (all p-values are better than p=1.9e-5, which is the p-value of the difference between days 4 and 10).
The entrance ring counts for all animals are independent (p=1.6e-34).

@@

The number of acini for day 4 was 18260~±~16555~acini for day 10 14359~±~5611~acini, for day 21 11203~±~2681~acini and for day 60 4277~±~777~acini.
These numbers were obtained by dividing the parenchymal volume of the lungs~(\protect\hyperlink{ref-wnl86DEM}{46}) with the Cavalieri-estimated volume of the acini.
To calculate these data we had to use data of two different studies.

@@

The total alveolar number for day 4 was found to be 0.64~±~0.34~million~acini, for day 10 1.23~±~0.39~million~acini, for day 21 2.65~±~0.57~million~acini, and for day 60 3.01~±~0.34~million~acini.
The total number of acini was estimated by dividing the mean parenchymal volume of the lungs~(\protect\hyperlink{ref-wnl86DEM}{46}) by the stereologically estimated acinar volume.
The number of alveoli for each animal is then calculated by multiplying the average entrance ring count per animal with the estimated number of acini per animal.

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung~(\protect\hyperlink{ref-14OP85b2F}{2}) and that information drawn from one lobe is representative of the whole lung~(\protect\hyperlink{ref-I9TmP6IU}{52}).
Furthermore, based on stereological data, Zeltner et al.~(\protect\hyperlink{ref-I9TmP6IU}{52}) were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy~(\protect\hyperlink{ref-Zv1XcgAO}{4}).
A first analysis of regional difference shows similar results as mentioned above.

@@

The mean acinar volume significantly increases during lung development from postnatal days 10 to 60 with all p-values better than 4.5e-6.
No significant difference was observed between postnatal days 4 and 10 (1.2× increase, p=0.08).
A large part of the acinus volume increase happens after day 21 (compare figure~\ref{fig:03}), which is consistent with previous literature~(\protect\hyperlink{ref-auNiE56G}{5}, \protect\hyperlink{ref-wnl86DEM}{46}).

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681} and is under revision in the \href{https://journals.physiology.org/journal/ajplung}{\emph{American Journal of Physiology - Lung Cellular and Molecular Physiology}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/d6fc17bd66bdca5cbff30d53e1dbc02d5351609b...eff43d9d3566f72329894e11e468eccd5bac3323}{has been updated} since the first round of review.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/eff43d9d3566f72329894e11e468eccd5bac3323/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/eff43d9d3566f72329894e11e468eccd5bac3323}{habi/acinar-analysis-manuscript@eff43d9}
on December 15, 2020.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(eff43d9)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(eff43d9)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Switzerland
\end{itemize}

@@

Lung development starts with the formation of the two lung buds.
During the embryonic stage (prenatal days E11--E13 in rats) the major airways and the pleura are formed.
During the pseudoglandular stage (E13--E18.5) most of the remaining airway generations are formed by branching morphogenesis.
Only few additional branches are added during the later stages (canalicular (E18.5--E20) and (E20 to postnatal day P4)) (\protect\hyperlink{ref-TsNwin2E}{34}).
Epithelial differentiation during the canalicular stage and angiogenetic activation of the capillaries lead to the first functional air-blood barriers in the lung (\protect\hyperlink{ref-ODQCVC23}{33}).
The saccular stage represents an intermediate stage where branching morphogenesis ceases and the developmental program switches to the phase of alveolarization.
During alveolarization, 90~\% of the adult gas-exchange area is formed by lifting off new septa from the existing gas-exchange surface (\protect\hyperlink{ref-auNiE56G}{5}, \protect\hyperlink{ref-1AhvLCPky}{19}, \protect\hyperlink{ref-XQLPrPTJ}{32}).
Alveolarization itself is divided into two distinct phases, the so-called classical (P4--P21) and continued alveolarization (P14 to approximately P60) (\protect\hyperlink{ref-d1YV8oOK}{35}, \protect\hyperlink{ref-wnl86DEM}{46}).
During classical alveolarization, new septa are formed starting from preexisting immature septa which contain a double-layered capillary network.
During continued alveolarization, new septa are formed from preexisting mature septa.
In general, during alveolarization the structure of the alveolar septa is changed to increase the efficiency of the lung (\protect\hyperlink{ref-d1YV8oOK}{35}, \protect\hyperlink{ref-wnl86DEM}{46}).

In this manuscript, the timing of lung development is given for rats while the same stages and developmental mechanisms are observed in every placentalian species studied so far, including humans.
However, the timing of lung development and especially the time point of birth relative to lung development is adapted in each species (\protect\hyperlink{ref-XQLPrPTJ}{32}, \protect\hyperlink{ref-ODQCVC23}{33}, \protect\hyperlink{ref-d1YV8oOK}{35}).

@@

Pulmonary acini represent the gas-exchanging units of the bronchial tree and are defined as the airways distal of the terminal bronchioles (\protect\hyperlink{ref-BbEFwEVD}{13}).
In humans, the acini contain approximately four generations of respiratory bronchioles before reaching the alveolar ducts.
The alveolar ducts start at the so-called bronchioalveolar duct junction (BADJ) where the lining of the inner airway surface abruptly changes from the cuboidal epithelium of the bronchioles into type 1 and type 2 alveolar epithelium which covers the alveoli.
The small tree of airways distal of the BADJ is called a ventilatory unit (\protect\hyperlink{ref-WDrIHn9p}{39}).
Since murine lungs do not possess respiratory bronchioles, the murine acini consist of one single ventilatory unit (\protect\hyperlink{ref-WDrIHn9p}{39}).
In this study we used rat lungs; hence we exclusively speak of acini.
To translate our results from rats to humans, monkeys, or dogs one would have to compare one rat acinus to one single ventilatory unit in these species.

Developmentally, the BADJ is of particular interest.
The junction is formed during the canalicular stage when the epithelia are differentiating.
The generation of the airway in which an individual murine acinus starts is defined during the canalicular stage and remains constant once it is formed.
As a consequence, in rats the number of formed acini also remains constant during the phase of alveolarization and thereafter (\protect\hyperlink{ref-14OP85b2F}{2}, \protect\hyperlink{ref-uFNlWogb}{3}).
The latter was somehow surprising because in rats the total lung volume increases by roughly a factor of 10 during alveolarization (\protect\hyperlink{ref-wnl86DEM}{46}).
Therefore, the mean volume of the acini has to increase by approximately the same factor.

The acinar architecture is important for ventilation, airflow (\protect\hyperlink{ref-CVl41LwO}{8}, \protect\hyperlink{ref-eb0gk6VO}{14}, \protect\hyperlink{ref-1DP2FRUSZ}{16}) and particle deposition (\protect\hyperlink{ref-18DcNLAv6}{22}, \protect\hyperlink{ref-yHHhvOtP}{36}, \protect\hyperlink{ref-wnl86DEM}{46}).

@@

On single two-dimensional sections---either histological or virtual from tomographic imaging---it is not possible to unambiguously detect which parts of the airways and alveoli are connected three-dimensionally.
To detect and delineate individual acini the full three-dimensional information is needed.
Historically, single or very few acini were delineated and extracted from manually traced serial sections (\protect\hyperlink{ref-c3aIB7f9}{28}, \protect\hyperlink{ref-CPhMy3Ie}{50}) or sectioned silicone casts (\protect\hyperlink{ref-1GJECnKBu}{30}) with considerable manual work.
Non-destructive three-dimensional imaging is best suited to acquire data sets that can be used to easily detect, delineate, and study large amounts of single acini.
This has been shown by Vasilescu et al.~(\protect\hyperlink{ref-C6OaY84D}{48}).
In their study, they reconstructed 22 mouse acini from tomographic data sets of four 12 weeks old mice (at least postnatal day 84, fully developed lungs) and compared those with scanning electron microscope images of acinar silicone rubber casts.
Similarly, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) analyzed the volume of 32 mouse acini from six animals, as well as the branching pattern of their internal airways.
All prior work known to us has analyzed considerably fewer acini than the 268 acini presented here and focuses only on a time point, at which the lungs of the animals are already fully developed while our data spans the postnatal days 4 to 60.

@@

While the global volume of the acini is well known and studied, only little is known about the range of the volume of individual acini throughout lung development.
We developed the hereby presented method to analyze large amounts of acini over the course of postnatal lung development in a time-efficient manner since no data on large amounts of acini is available up to now.
Our aim was to understand the size distribution of the acini during lung development including the contribution of the alveoli.
The observed range \emph{de}creased during postnatal lung development by a factor of 6--7, which represents a completely unexpected result.
The obtained data are essential for further investigations, like the influence of the size of the acini on ventilation or gas-washout, respectively.
E.g., we are currently using our computational fluid dynamics model to simulate exactly this influence (own unpublished results and (\protect\hyperlink{ref-eb0gk6VO}{14})).

@@

In the present study, a superset of the animals described by Haberthür et al.~(\protect\hyperlink{ref-7YLeeyu}{11}) were used.
Tschanz et al.~(\protect\hyperlink{ref-wnl86DEM}{46}) described a biphasic formation of new alveoli on the same set of animals.
The stereological analysis presented here represents a part of the 3R-initiative (replacement, reduction, and refinement) (\protect\hyperlink{ref-199ALtdJt}{53}) for the ethical use and the reduction of the number of animals sacrificed for science.

Briefly, we extracted the lungs of Sprague-Dawley rats after having deeply anesthetized them with a mixture of medetomidine, midazolam, and fentanyl (\protect\hyperlink{ref-IdRLVLjl}{7}).
The rats were euthanized by exsanguination during the removal of the lung.
The lungs were fixed with 2.5~\% glutaraldehyde by intratracheal instillation and kept under constant pressure (20 cm H\textsubscript{2}O) during fixation to prevent recoiling of the tissue.
The samples were prepared for tomographic imaging by post-fixation with 1~\% osmium tetroxide and staining with 4~\% uranyl acetate, as well as by dehydration in a graded series of ethanol and embedding in paraffin.
In total we assessed 12 animals, on postnatal days P4, P10, P21, and P60 (N=3 per day.
From now on, we will omit the postnatal prefix and only mention the day.).

@@

Since our samples were substantially larger than the field of view of the detector, we applied the so-called wide-field scanning method (\protect\hyperlink{ref-VELl7OiR}{12}) to increase the field of view horizontally.
For each sub-scan, we recorded 3578 projections and laterally stitched them so that their combined field of view covered the whole sample width.
Additionally, two or three such wide-field scans were stacked vertically to match the sample height.

@@

The stereological estimation of the alveolar number was performed according to the standards for quantitative assessment of lung structure from the \href{http://www.thoracic.org/}{American Thoracic Society} (\protect\hyperlink{ref-dNc8FfNn}{17}) to guarantee accurate and unbiased results.
As a basis for the estimation the volumes of the lung lobes were first determined by water displacement (\protect\hyperlink{ref-KGbSQovR}{31}).
The lobe-volume was estimated a second time by the Cavalieri method.
Both volumes were used to calculate the shrinkage, which was used for the correction of the volume data (\protect\hyperlink{ref-wnl86DEM}{46}).

To perform the stereological analysis, each DICOM file from the isolated acini was sampled into a systematic random set of single images using a MATLAB script (\protect\hyperlink{ref-12Z2YPzm8}{10}).
The stereological assessment was performed with the \href{http://stepanizer.com/}{STEPanizer} (\protect\hyperlink{ref-nPoQ2EIB}{45}), an easy-to-use software tool for the stereological assessment of digital images.
Details of the stereological assessment have previously been described (\protect\hyperlink{ref-7YLeeyu}{11}).
Briefly, we counted the appearance or disappearance of alveolar wall interruptions.
Under the assumption that these only occur in the region of the alveolar mouth opening and correspond to alveolar entrance rings (\protect\hyperlink{ref-QiAxY2i3}{18}) we counted said interruptions on paired images spanning the whole acinus volume.
By using the disector method (\protect\hyperlink{ref-FJ9FoB4m}{38}) we were thus able to estimate the number of alveoli in 268 of the 701 exported acini.
The acini we analyzed were subsampled from the full data by systematic uniform random sampling (\protect\hyperlink{ref-dNc8FfNn}{17}).
Briefly, a defined fraction of the acini of one animal were analyzed, e.g.~one third.
During segmentation, the acini were numbered.
Based on these numbers every third acinus was selected, randomly starting with the first, second, or third one.

Previously, we have shown that the stereological assessment of the acinar volume (according to the Cavalieri principle (\protect\hyperlink{ref-FE9HLB4f}{6})) gives comparable results to an automated volume extraction by counting the segmented pixels in the tomographic data sets (\protect\hyperlink{ref-7YLeeyu}{11}).
Due to variations in the gray value, the automatic segmentation underestimates the volume of the single acini.
The best-suited approach to obtain unbiased results is to assess the volume manually according to the Cavalieri principle, which is what we did for the 268 acini presented in this manuscript.
The sets of JPG slices and the raw results from the stereological analysis with the STEPanizer are available on request.

@@

All the stereologically assessed data was processed in a \href{http://jupyter.org/}{Jupyter} (\protect\hyperlink{ref-pQ6Wbz73}{21}) notebook, producing all the results, including the statistical data and plots shown below.
The notebook with all its history is freely available on GitHub (\protect\hyperlink{ref-V87xbt0b}{9}).
The performed calculations are described in detail in the \protect\hyperlink{results}{Results} section.
In the plots, semitransparent circles mark the single observations.
The box shows the 25~\%--75~\% quartiles range of the data.
The whiskers extend to the rest of the distribution.
Outliers are determined as a function of the inter-quartile range and are shown outside the whiskers.
Numerical values in the text are given as averages ± standard deviation.
P-values in the text and figure legends are given as precise numbers, as suggested by Amhrein et al.~(\protect\hyperlink{ref-o21zxPIu}{1}).

Usually, we performed a Shapiro-Wilk test for normality (\protect\hyperlink{ref-IkHrgIj3}{37}) to test whether we can use an U-test for assessing the differences, namely a two-sided Mann-Whitney rank test (\protect\hyperlink{ref-MyzxBJ57}{25}).
This rank test was used to assess the differences between the possible combinations.
An additional Kruskal-Wallis H-test for independent samples (\protect\hyperlink{ref-ON1Bppkk}{23}) was used to test for sample independence.
The statistical analysis was performed in the aforementioned notebook by using the statistical functions of the Scientific Computing Tools for Python (\protect\hyperlink{ref-8Miti2Gz}{49}) or GraphPad Prism 7.01 (GraphPad Software, San Diego, CA, USA).

@@

The alveolar number shows highly significant differences Šidák-corrected p-value \textless{} 0.00167 (\protect\hyperlink{ref-GtCynqsI}{44})) between all possible combination of days.
All possible combinations of entrance ring counts per day are significantly different (all p-values are better than p=1.9e-5, which is the p-value of the difference between days 4 and 10).
The entrance ring counts for all animals are independent (p=1.6e-34).

@@

The number of acini for day 4 was 18260~±~16555~acini for day 10 14359~±~5611~acini, for day 21 11203~±~2681~acini and for day 60 4277~±~777~acini.
These numbers were obtained by dividing the parenchymal volume of the lungs (\protect\hyperlink{ref-wnl86DEM}{46}) with the Cavalieri-estimated volume of the acini.
To calculate these data we had to use data of two different studies.

@@

The total alveolar number for day 4 was found to be 0.64~±~0.34~million~acini, for day 10 1.23~±~0.39~million~acini, for day 21 2.65~±~0.57~million~acini, and for day 60 3.01~±~0.34~million~acini.
The total number of acini was estimated by dividing the mean parenchymal volume of the lungs (\protect\hyperlink{ref-wnl86DEM}{46}) by the stereologically estimated acinar volume.
The number of alveoli for each animal is then calculated by multiplying the average entrance ring count per animal with the estimated number of acini per animal.

@@

We can easily estimate the total number of acini during lung development from our data.
The limitations of this study are mainly twofold: a) due to technical reasons (i.e.~available beamtime and limitations of the setup we could not image the entire lung and b) lung location (i.e.~lower medial tip of the right lobe).
However, our sample is more than suitable for the drawn conclusions because it was previously shown that the right lower lobe is a valid sample for the entire lung (\protect\hyperlink{ref-14OP85b2F}{2}) and that information drawn from one lobe is representative of the whole lung (\protect\hyperlink{ref-I9TmP6IU}{52}).
Furthermore, based on stereological data, Zeltner et al.~(\protect\hyperlink{ref-I9TmP6IU}{52}) were able to show that the lung parenchyma is quite homogeneous.
Looking at individual acini, Kizhakke Puliyakote et al.~(\protect\hyperlink{ref-RGBeCf8v}{20}) observed that the peripheral, pleura facing acini possess an acinar volume which is two thirds larger than the volume of the central ones.
However, they did not report additional regional differences, e.g.~between acini located cranially and caudally.
Own preliminary data point to the same result.
We imaged an entire rat lung \emph{in situ} immediately post mortem using high-resolution synchrotron radiation based X-ray tomographic microscopy (\protect\hyperlink{ref-Zv1XcgAO}{4}).
A first analysis of regional difference shows similar results as mentioned above.

@@

The mean acinar volume significantly increases during lung development from postnatal days 10 to 60 with all p-values better than 4.5e-6.
No significant difference was observed between postnatal days 4 and 10 (1.2× increase, p=0.08).
A large part of the acinus volume increase happens after day 21 (compare figure~\ref{fig:03}), which is consistent with previous literature (\protect\hyperlink{ref-auNiE56G}{5}, \protect\hyperlink{ref-wnl86DEM}{46}).

As mentioned in the \protect\hyperlink{introduction}{introduction} section, Barré et al.~(\protect\hyperlink{ref-14OP85b2F}{2}) have shown that the number of acini remains constant during lung development, albeit in a different strain of animals (Wistar Bern compared to Sprague-Dawley in this study).
The total volume of the alveolar space increases by roughly a factor of 11 during alveolarization (\protect\hyperlink{ref-wnl86DEM}{46}).
Therefore, we expect the mean volume of the acini to increase by at least the same factor.
